\LARGE \textbf{Constrained Geometric Attitude Control on \( \SO \)}

### 2016-11-30 manuscript.tex
Bigger figures which span both columns

\Cref{eqn:psi} is composed of an attractive term, \( A (R) \) toward the desired attitude, and a repulsive term, \( B(R) \) away from the undesired direction \( R^T v \).
In order to visualize the attitude error function on \( \SO \), we utilize a spherical coordinate representation.
Recall, that the spherical coordinate system represents the position of a point relative to an origin in terms of a radial distance, azimuth, and elevation.
This coordinate system is commonly used to define locations on the Earth in terms of a latitude and longitude.
Similarly, the positions of celestial objects are defined on the celestial sphere in terms of right ascension and declination. 
\begin{figure}[H]%
    \centering 
    \subfigure[Attractive \( A(R) \) \label{fig:attract_error}]{\includegraphics[width=0.45\columnwidth]{attract_error} }~
    \subfigure[Repulsive \( B(R) \) \label{fig:avoid_error} ]{\includegraphics[width=0.45\columnwidth]{avoid_error} }\\%add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
    \subfigure[Configuration \( \Psi \) \label{fig:combined_error}]{\includegraphics[width=0.45\columnwidth]{combined_error} }%
    \caption{Configuration error function visualization}
    \label{fig:config_error} 
\end{figure}%
We apply this concept and parametrize the rotation matrix \( R \in \SO \) in terms of the spherical angles \( \SI{-180}{\degree} \leq \lambda \leq \SI{180}{\degree}  \) and \( \SI{-90}{\degree} \leq \beta \leq \SI{90}{\degree} \). 
Using the elementary Euler rotations, the rotation matrix is now defined as \( R = \exp( \lambda \hat{e}_2) \exp( \beta \hat{e}_3) \).
We iterate over the domains of \( \lambda\) and \(\beta\) in order to rotate the body-fixed vector \( r \) throughout the two-sphere \( \S^2 \).
Applying this method,~\cref{fig:config_error} allows us to visualize the error function on \( \SO \).
The horizontal axes of~\cref{fig:config_error} represent the domain of the spherical angles \( \lambda \) and \( \beta \) in degrees, while the vertical axes represent the unitless magnitude of the error functions defined in~\cref{eqn:psi,eqn:A,eqn:B}.
The attractive error function, given by~\cref{eqn:A}, has been previously used for attitude control on \(\SO\).
The potential well of \( A(R)\) is illustrated in~\cref{fig:attract_error}, where the desired attitude lies at the minimum of \( A(R) \).

@@

The initial state is defined as \(R_0 =  \exp(\ang{225} \times \frac{\pi}{180} \hat{e}_3), \Omega_0 = 0\). 
The desired state is \( R_d = I,\Omega_d = 0\).
\begin{figure}[H] 
    \centering 
    \subfigure[Attitude error vector \(e_R\) \label{fig:eR_con}]{\includegraphics[width=0.5\columnwidth]{eR_con}  }~
    \subfigure[Configuration error \( \Psi \) \label{fig:Psi_con}]{\includegraphics[width=0.5\columnwidth]{figures/Psi_con.pdf} }
    \subfigure[Angle to constraints \label{fig:con_angles_con}]{\includegraphics[width=0.5\columnwidth]{figures/constraint_angles_con.pdf} }  
    \caption{Attitude stabilization without adaptive update law}
    \label{fig:con} 
\end{figure}
We show simulation results for the system stabilizing about the desired attitude with and without the adaptive update law from~\Cref{prop:adaptive_control}.
We assume a fixed disturbance of \(\Delta = \begin{bmatrix} 0.2 & 0.2 & 0.2 \end{bmatrix}^T \si{\newton\meter}\), with the function \( W(R,\Omega) = I \).
This form is equivalent to an integral control term which penalizes deviations from the desired configuration.
The first term of~\cref{eqn:delta_dot} has the effect of increasing the proportional gain of the control system, since the time derivative of the attitude error vector, \( \dot{e}_{R} \), is linear with respect to the angular velocity error vector \( e_\Omega\).
\begin{figure}[H] 
	\centering 
    \subfigure[Configuration error \( \Psi \)\label{fig:Psi_adapt}]{\includegraphics[width=0.5\columnwidth]{Psi_adapt} }~
    \subfigure[Angle to constraints \label{fig:con_angles}]{\includegraphics[width=0.5\columnwidth]{constraint_angles} }\\
    \subfigure[Disturbance estimate \( \bar \Delta \)  \label{fig:delta_adapt} ]{\includegraphics[width=0.5\columnwidth]{Delta_adapt}}~
    \subfigure[Attitude trajectory \label{fig:cad_adapt}]{\includegraphics[trim=9cm 2.5cm 0 0,clip,width=0.645\columnwidth]{cad_adapt} }
	\caption{Attitude stabilization with adaptive update law}
	\label{fig:adapt} 
\end{figure}

Simulation results without the adaptive update law are shown in~\cref{fig:con}.
\Cref{fig:eR_con} shows each component of the attitude error vector,~\cref{eqn:eR}, over the simulation time span.
\Cref{fig:Psi_con} shows the  magnitude of the combined error function,~\cref{eqn:psi}.
Without the update law, the system does not achieve zero steady state error. 
\Cref{fig:Psi_con} shows that the configuration error function does not converge to zero and there exist steady state errors.
In spite of the uncompensated disturbance, the system is able to avoid the constrained regions as shown in~\cref{fig:con_angles_con}.
The angle to each of the constraints, which is measured in degrees and given by \( \arccos(r^T R^T v_i) \), is always greater than the specified angle, \( \theta_i \), in~\Cref{tab:constraints}.

@@

A sensor pointing direction is defined in the body-fixed frame of the hexrotor as \( r = [1,0,0]^T \).
We define an obstacle in the inertial frame as \( v = [\frac{1}{\sqrt{2}}, \frac{1}{\sqrt{2}}, 0]^T \) with \( \theta = \ang{12} \).
An initial state is defined as \(R(0) = \exp( \frac{\pi}{2} \hat{e}_3) \), while the desired state is \(R_d =I \).
This results in the UAV performing a \ang{90} yaw rotation about the vertical axis of the spherical joint and the constrained region is on the shortest path connecting $R_0$ and $R_d$. 
The attitude control system is identical to the one presented in~\Cref{prop:adaptive_control} with the exception of a gravity moment term and the following parameters: \(k_R = 0.4, k_\Omega = 0.7 ,c = 0.1 , \alpha = 8 \text{ and } k_\Delta = 0.05\).
The parameters are modified to account for the differences in the hardware model of the hexrotor.
\begin{figure}[H]
	\centering 
    \subfigure[{Attitude error vector \(e_R\)} \label{fig:eR_exp}]{\includegraphics[width=0.5\columnwidth]{eR_exp}  }~
	\subfigure[{Configuration error \( \Psi \)} \label{fig:Psi_exp} ]{\includegraphics[width=0.5\columnwidth]{Psi_exp} }\\
	\subfigure[{Control input \( u\) }\label{fig:u_exp}]{\includegraphics[width=0.5\columnwidth]{u_exp}  }~
	\subfigure[{Attitude Trajectory } \label{fig:traj_exp}]{\includegraphics[width=0.5\columnwidth]{traj_exp} }\\
    \subfigure[{Angle to constraint } \label{fig:con_angle_exp}]{\includegraphics[width=0.5\columnwidth]{figures/constraint_angles_exp.pdf} }
	\caption{Constrained Attitude stabilization experiment}
	\label{fig:exp} 
\end{figure}
The experimental results are shown in~\Cref{fig:exp}.
\Cref{fig:eR_exp} shows the behavior of each of the components of the attitude error vector, defined by~\cref{eqn:eR}, over the experiment time span.
\Cref{fig:Psi_exp} shows the time history of the attitude error function, defined by~\cref{eqn:psi}.
\Cref{fig:u_exp} shows the magnitude of each component of the control input in \si{\newton\meter}, which is computed from~\cref{eqn:adaptive_control}.
Finally,~\cref{fig:con_angle_exp} shows the angle between the body-fixed sensor and the obstacle in degrees.
In order to maneuver the system ``close" to the constrained zone we utilize several intermediary set points on either side of the obstacle.
From the initial attitude the hexrotor rotates to the first set point, pauses, and then continues around the obstacle to the second set point before continuing toward the desired attitude.
As a result this creates the stepped behavior of the configuration error history as shown in~\cref{fig:Psi_exp}.

>>>>

\Cref{eqn:psi} is composed of an attractive term, \( A (R) \) toward the desired attitude, and a repulsive term, \( B(R) \) away from the undesired direction \( R^T v \).
In order to visualize the attitude error function on \( \SO \), we utilize a spherical coordinate representation.
Recall, that the spherical coordinate system represents the position of a point relative to an origin in terms of a radial distance, azimuth, and elevation.
This coordinate system is commonly used to define locations on the Earth in terms of a latitude and longitude.
Similarly, the positions of celestial objects are defined on the celestial sphere in terms of right ascension and declination. 
\end{multicols}
\begin{figure}[H]%
    \centering 
    \subfigure[Attractive \( A(R) \) \label{fig:attract_error}]{\includegraphics[width=0.3\columnwidth]{attract_error} }~
    \subfigure[Repulsive \( B(R) \) \label{fig:avoid_error} ]{\includegraphics[width=0.3\columnwidth]{avoid_error} }~
    \subfigure[Combined \( \Psi \) \label{fig:combined_error}]{\includegraphics[width=0.3\columnwidth]{combined_error} }%
    \caption{Visualization of Configuration Error Functions using spherical coordinate representation}
    \label{fig:config_error} 
\end{figure}%
\begin{multicols}{2}
We apply this concept and parametrize the rotation matrix \( R \in \SO \) in terms of the spherical angles \( \SI{-180}{\degree} \leq \lambda \leq \SI{180}{\degree}  \) and \( \SI{-90}{\degree} \leq \beta \leq \SI{90}{\degree} \). 
Using the elementary Euler rotations, the rotation matrix is now defined as \( R = \exp( \lambda \hat{e}_2) \exp( \beta \hat{e}_3) \).
We iterate over the domains of \( \lambda\) and \(\beta\) in order to rotate the body-fixed vector \( r \) throughout the two-sphere \( \S^2 \).
Applying this method,~\cref{fig:config_error} allows us to visualize the error function on \( \SO \).
The horizontal axes of~\cref{fig:config_error} represent the domain of the spherical angles \( \lambda \) and \( \beta \) in degrees, while the vertical axes represent the unitless magnitude of the error functions defined in~\cref{eqn:psi,eqn:A,eqn:B}.
The attractive error function, given by~\cref{eqn:A}, has been previously used for attitude control on \(\SO\).
The potential well of \( A(R)\) is illustrated in~\cref{fig:attract_error}, where the desired attitude lies at the minimum of \( A(R) \).

@@

The initial state is defined as \(R_0 =  \exp(\ang{225} \times \frac{\pi}{180} \hat{e}_3), \Omega_0 = 0\). 
The desired state is \( R_d = I,\Omega_d = 0\).
\end{multicols}
\begin{figure}
    \centering 
    \subfigure[Attitude error vector \(e_R\) components \label{fig:eR_con}]{\includegraphics[width=0.3\columnwidth]{eR_con}  }~
    \subfigure[Configuration error \( \Psi \) \label{fig:Psi_con}]{\includegraphics[width=0.3\columnwidth]{figures/Psi_con.pdf} }
    \subfigure[Angle to each constraint \label{fig:con_angles_con}]{\includegraphics[width=0.3\columnwidth]{figures/constraint_angles_con.pdf} } 
    \caption{Attitude stabilization without adaptive update law}
    \label{fig:con} 
\end{figure}
\begin{multicols}{2}
We show simulation results for the system stabilizing about the desired attitude with and without the adaptive update law from~\Cref{prop:adaptive_control}.
We assume a fixed disturbance of \(\Delta = \begin{bmatrix} 0.2 & 0.2 & 0.2 \end{bmatrix}^T \si{\newton\meter}\), with the function \( W(R,\Omega) = I \).
This form is equivalent to an integral control term which penalizes deviations from the desired configuration.
The first term of~\cref{eqn:delta_dot} has the effect of increasing the proportional gain of the control system, since the time derivative of the attitude error vector, \( \dot{e}_{R} \), is linear with respect to the angular velocity error vector \( e_\Omega\).
\end{multicols}
\begin{figure} 
	\centering 
    \subfigure[Configuration error \( \Psi \)\label{fig:Psi_adapt}]{\includegraphics[width=0.4\columnwidth]{Psi_adapt} }~
    \subfigure[Angle to each constraint \label{fig:con_angles}]{\includegraphics[width=0.4\columnwidth]{constraint_angles} }\\
    \subfigure[Disturbance estimate \( \bar \Delta \) components \label{fig:delta_adapt} ]{\includegraphics[width=0.4\columnwidth]{Delta_adapt}}~
    \subfigure[Attitude trajectory \label{fig:cad_adapt}]{\includegraphics[trim={10cm 0 10cm 0},clip,width=0.4\columnwidth]{cad_adapt} }
	\caption{Attitude stabilization with adaptive update law}
	\label{fig:adapt} 
\end{figure}
\begin{multicols}{2}
Simulation results without the adaptive update law are shown in~\cref{fig:con}.
\Cref{fig:eR_con} shows each component of the attitude error vector,~\cref{eqn:eR}, over the simulation time span.
\Cref{fig:Psi_con} shows the  magnitude of the combined error function,~\cref{eqn:psi}.
Without the update law, the system does not achieve zero steady state error. 
\Cref{fig:Psi_con} shows that the configuration error function does not converge to zero and there exist steady state errors.
In spite of the uncompensated disturbance, the system is able to avoid the constrained regions as shown in~\cref{fig:con_angles_con}.
The angle to each of the constraints, which is measured in degrees and given by \( \arccos(r^T R^T v_i) \), is always greater than the specified angle, \( \theta_i \), in~\Cref{tab:constraints}.

@@

A sensor pointing direction is defined in the body-fixed frame of the hexrotor as \( r = [1,0,0]^T \).
We define an obstacle in the inertial frame as \( v = [\frac{1}{\sqrt{2}}, \frac{1}{\sqrt{2}}, 0]^T \) with \( \theta = \ang{12} \).
An initial state is defined as \(R(0) = \exp( \frac{\pi}{2} \hat{e}_3) \), while the desired state is \(R_d =I \).
This results in the UAV performing a \ang{90} yaw rotation about the vertical axis of the spherical joint and the constrained region is on the shortest path connecting $R_0$ and $R_d$. 
The attitude control system is identical to the one presented in~\Cref{prop:adaptive_control} with the exception of a gravity moment term and the following parameters: \(k_R = 0.4, k_\Omega = 0.7 ,c = 0.1 , \alpha = 8 \text{ and } k_\Delta = 0.05\).
The parameters are modified to account for the differences in the hardware model of the hexrotor.
\end{multicols}
\begin{figure}
	\centering 
    \subfigure[{Attitude error vector \(e_R\)} components\label{fig:eR_exp}]{\includegraphics[width=0.3\columnwidth]{eR_exp}  }~
	\subfigure[{Configuration error \( \Psi \)} \label{fig:Psi_exp} ]{\includegraphics[width=0.3\columnwidth]{Psi_exp} }~
	\subfigure[{Control input \( u\) } components\label{fig:u_exp}]{\includegraphics[width=0.3\columnwidth]{u_exp}  }\\
	\subfigure[{Attitude Trajectory } \label{fig:traj_exp}]{\includegraphics[width=0.3\columnwidth]{traj_exp} }~
    \subfigure[{Angle to constraint } \label{fig:con_angle_exp}]{\includegraphics[width=0.3\columnwidth]{figures/constraint_angles_exp.pdf} }
	\caption{Constrained Attitude stabilization experiment}
	\label{fig:exp} 
\end{figure}
\begin{multicols}{2}
The experimental results are shown in~\Cref{fig:exp}.
\Cref{fig:eR_exp} shows the behavior of each of the components of the attitude error vector, defined by~\cref{eqn:eR}, over the experiment time span.
\Cref{fig:Psi_exp} shows the time history of the attitude error function, defined by~\cref{eqn:psi}.
\Cref{fig:u_exp} shows the magnitude of each component of the control input in \si{\newton\meter}, which is computed from~\cref{eqn:adaptive_control}.
Finally,~\cref{fig:con_angle_exp} shows the angle between the body-fixed sensor and the obstacle in degrees.
In order to maneuver the system ``close" to the constrained zone we utilize several intermediary set points on either side of the obstacle.
From the initial attitude the hexrotor rotates to the first set point, pauses, and then continues around the obstacle to the second set point before continuing toward the desired attitude.
As a result this creates the stepped behavior of the configuration error history as shown in~\cref{fig:Psi_exp}.
